---
title: "COPD Statistical Analysis"
format:
  html:
    toc: true
    code-fold: true
    code-summary: "Show R code"
    code-copy: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup

library(tidyverse)
library(broom)
library(janitor)
library(here)
library(scales)
library(gt)

source(here("R", "01_data_cleaning.R"))
source(here("R", "02_exploratory_analysis.R"))
source(here("R", "03_statistical_models.R"))
source(here("R", "04_visualizations.R"))
```

## Data Import

The project expects a de-identified CSV file at `data/copd_data.csv`.

```{r}
#| label: import-data

copd_raw <- read_csv(here("data", "copd_data.csv"), show_col_types = FALSE)
copd_data <- clean_copd_data(copd_raw)
```

## Study Population

```{r}
#| label: participant-summary

participant_summary(copd_data) |>
  gt()
```

## COPD Prevalence by Smoking Status

```{r}
#| label: fig-smoking-prevalence
#| fig-cap: "COPD prevalence by smoking status"

plot_copd_by_smoking(copd_data)
```

## COPD Prevalence by Age Group

```{r}
#| label: fig-age-prevalence
#| fig-cap: "COPD prevalence by age group"

plot_copd_by_age(copd_data)
```

## Adjusted Logistic Regression

The primary model estimates the association between COPD and age, sex, smoking status, and body mass index.

```{r}
#| label: fit-model

copd_model <- fit_copd_model(copd_data)

model_odds_ratios(copd_model) |>
  gt()
```

## Interpretation

Replace this section with your findings. Focus on effect size, confidence intervals, practical importance, and study limitations rather than only reporting p-values.

## Limitations

Potential limitations may include missing data, self-reported smoking history, cross-sectional design, residual confounding, sample selection, and limited generalizability.
